---
title       : Singapore Public Housing Resale Prices
subtitle    : Developing Data Products Assignment 
author      : by Chan Chee-Foong for Data Science Specialisation by Johns Hopkins University
job         : 
framework   : revealjs        # {io2012, html5slides, shower, dzslides, deckjs, revealjs, impressjs...}
revealjs    : {theme: solarized, transition: cube}
# highlighter : highlight.js  # {highlight.js, prettify, highlight}
# hitheme     : tomorrow      # 
# widgets     : []            # {mathjax, quiz, bootstrap}
ext_widgets : {rCharts: libraries/nvd3} 
mode        : selfcontained # {standalone, draft}
knit        : slidify::knit2slides
---


``` {r initialise, echo = FALSE, results = "hide", warning = FALSE, message = FALSE}
library(ggplot2)
library(lubridate)
library(dplyr)
library(reshape2)
library(rCharts)
library(knitr)

datadir <- "./data"
datafile <- "resale-flat-prices-based-on-registration-date-from-march-2012-onwards.csv"
zipfile <- "resale-flat-prices.zip"

datadirfile <- paste(datadir, datafile, sep="/")
zipdirfile <- paste(datadir, zipfile, sep="/")

if (!file.exists(datadirfile)) {
    dir.create(datadir)
    url <- "https://data.gov.sg/dataset/7a339d20-3c57-4b11-a695-9348adfd7614/download"
    download.file(url, destfile = zipdirfile, mode = 'wb')
    unzip (zipdirfile, exdir = datadir)
}

txnData <- read.csv(datadirfile)

txnData$size <- sapply(strsplit(as.character(txnData$flat_type)," "), function(str) str[[1]])
txnData[!(txnData$size %in% c(1:5)),'size'] <- 6
txnData$size <- as.numeric(txnData$size)

Data <- txnData %>% group_by(month, flat_type) %>% summarise(avg_price = mean(resale_price), number_of_sales = n())

oData <- txnData %>% group_by(town) %>% summarise(countTown = n()) %>% arrange(desc(countTown))
txnData$town <- factor(txnData$town, levels = oData$town)

townData <- txnData %>% group_by(town, flat_type) %>% summarise(avg_price = mean(resale_price), number_of_sales = n())

Data$date <- ymd(paste0(Data$month,'-01'))
Data$date <- as.Date(Data$date)
Data$year <- year(Data$date)


# For Slide 5 - PSF calculation
sqm2sqft <- 10.7639

txnData$floor_area_sqft <- sqm2sqft * txnData$floor_area_sqm
txnData$psf <- txnData$resale_price/txnData$floor_area_sqft

psfData <- txnData %>% group_by(town) %>% summarise(avg_psf = mean(psf)) %>% arrange(desc(avg_psf))
psfData$town <- factor(psfData$town, levels = psfData$town)

psfData$region <- 'CENTRAL'
psfData[psfData$town %in% c('CENTRAL AREA'),]$region <- 'DOWNTOWN'
psfData[psfData$town %in% c('BEDOK','PASIR RIS','TAMPINES'),]$region <- 'EAST'
psfData[psfData$town %in% c('SEMBAWANG','WOODLANDS','YISHUN'),]$region <- 'NORTH'
psfData[psfData$town %in% c('ANG MO KIO','HOUGANG','PUNGGOL','SENGKANG','SERANGOON'),]$region <- 'NORTH-EAST'
psfData[psfData$town %in% c('BUKIT BATOK','BUKIT PANJANG','CHOA CHU KANG','CLEMENTI','JURONG EAST','JURONG WEST'),]$region <- 'WEST'

```

### Singapore Public Housing Resale Prices
  
<br>  
  
by Chan Chee-Foong  
<br>
on 4 Aug 2016  
<br>
for Developing Data Products Assignment  
<br>
of the Data Science Specialisation  
<br>
by Johns Hopkins University  


--- .class #id 

## Exploratory Data Analysis

1. Data (Resale Flat Prices) is downloaded directly from http://data.gov.sg.
2. The public housing resale transacted prices are provided by Housing 
Development Board in Singapore.
3. This analysis uses all available data in the downloaded data size.  The towns are further classified into regions based on information found on this [website here] (https://data.gov.sg/dataset/master-plan-2014-planning-area-boundary-no-sea).
4. A more interactive application (ShinyApps) developed to explore the data can be found 
[here] (https://edgetrader.shinyapps.io/DevelopingDataProductsAssignment/).
5. All the charts provided in this presentation are generated using R ggplot emebedded in this slidify document.

--- .class #id 

### Total Number of Monthly Sales

Lowest number of transaction recorded on Feb 2013 due to government cooling measures on 12 Jan 2013.

```{r Slide3, results = 'asis', echo = FALSE, warning = FALSE, message = FALSE, fig.height=6, fig.width=12}

ggplot(data=Data, aes(x=month, y=number_of_sales, fill = flat_type)) +
    theme(legend.position="top") +
    geom_bar(stat="identity") +
    scale_fill_discrete(name = "FLAT TYPES") +
    xlab("MONTHS") + 
    ylab("NUMBER OF TRANSACTIONS") +
    theme(legend.position="top", legend.title = element_text(face = "bold")) +
    guides(fill=guide_legend(nrow=1)) +
    theme(axis.title.x = element_text(size=12),
        axis.text.x  = element_text(angle=90, vjust=0.5, size=8))

```


--- .class #id 

### Total of Sales Transactions by Town

Top 3 towns having the most sales transactions:  
Jurong West, Woodlands and Tampines

```{r Slide4, results = 'asis', echo = FALSE, warning = FALSE, message = FALSE, fig.height=6, fig.width=12}

ggplot(data=townData, aes(x=town, y=number_of_sales, fill=flat_type)) +
    theme(legend.position="top") +
    geom_bar(stat="identity") +
    scale_fill_discrete(name = "FLAT TYPES") +
    xlab("TOWNS") + 
    ylab("NUMBER OF TRANSACTIONS") +
    theme(legend.position="top", legend.title = element_text(face = "bold")) +
    guides(fill=guide_legend(nrow=1)) +
    theme(axis.title.x = element_text(size=12),
        axis.text.x  = element_text(angle=90, vjust=0.5, size=8))

```

--- .class #id 

### Price per sqft in various regions and towns

Most expensive public housing flats are found in towns close to city centre.

```{r Slide5, results = 'asis', echo = FALSE, warning = FALSE, message = FALSE, fig.height=6, fig.width=12}

ggplot(data=psfData, aes(x=town, y=avg_psf, fill = region)) +
    geom_bar(stat="identity", colour = 'black') +
    coord_cartesian(ylim = c(350,700)) +
    geom_text(aes(label=round(avg_psf,0)), vjust=-.5) +
    scale_fill_discrete(name = "REGIONS") +
    xlab("TOWNS") + 
    ylab("AVERAGE PRICE PER SQFT (S$)") +
    theme_classic() +
    theme(legend.position="top", legend.title = element_text(face = "bold")) +
    guides(fill=guide_legend(nrow=1)) +
    theme(axis.title.x = element_text(size=12),
          axis.text.x  = element_text(angle=90, vjust=0.5, size=8))

```